Fishery Markets and Large-Scale Marine Conservation

### 2019-05-30 docs/Markets_Conservation_May30.tex
Added all files

\title{Marine Conservation In Fishing Effort Markets}

@@

\author{Juan CarlosVillase\~{n}or-Derbez,$^{1\ast}$ John Lynham,$^{2}$ Christopher Costello$^{1}$\\
\\
\normalsize{$^{1}$Bren School of Environmental Science \& Management,}\\
\normalsize{University of California at Santa Barbara, Santa Barbara, CA}\\
\normalsize{$^{2}$Department of Economics, University of Hawaii at Manoa, Honolulu, HI}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail: juancarlos@ucsb.edu.}
}

@@

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}
Marine Protected Areas have the primary objective of conserving bounded waters by displacing or limiting fishing effort. With global conservation targets seeking to protect 10\% of the world’s ocean by 2020, it is important that the spatial displacement of economic activities is fully considered. A spatial closure in a trans-boundary fishery can potentially result in costs to the implementing country and benefits to neighboring countries. We show that effort markets may reduce up to 70\% of the costs of Marine Protected Areas, and that allocation rules based on biomass (rather than effort) may further incentivize conservation. We then use identification of fishing activity via Automatic Identification Systems to evaluate the effect that the Phoenix Islands Protected Area (PIPA, 397,447 km\textsuperscript{2}, implemented in 2015 in Kiribati) had on vessel distribution and behavior. Aggregate fishing effort remains relatively constant following the closure, but is significantly reduced within Kiribati, especially by vessels that historically fished within PIPA before its implementation. We find evidence that Kiribati may be trading its surplus of vessel-days, which would significantly reduce the costs of displaced fishing effort. We use these insights to predict the implications of a proposed Large-Scale Marine Protected Area (LSMPA) in Palau (another PNA member) and estimate that effort markets could reduce the cost of implementation by as much as 77\%. Our work highlights the role of effort markets in enabling conservation, and the importance of allocation rules that praise, not punish, conservation. The Parties to the Nauru Agreement have successfully developed a management scheme that allows for conservation interventions to come at a low cost, and could be a model for trans-boundary conservation across the globe.
\end{sciabstract}

@@

\section{Introduction}\label{introduction}

Humans are increasingly utilizing the oceans. Multiple ocean uses such as off-shore aquaculture, conservation, energy harvesting, deep-sea mining, and fisheries are likely to compete for space. As we move forward with blue growth, we must understand the potential effects of activities displacing each other and establish causal links between past management interventions and their outcomes \cite{burgess_2018}. One of the most notable spatial interventions is the creation of no-take Marine Protected Areas (MPAs), which seek to conserve the environment by eliminating fishing effort within their waters.

Today, a small number of Large-scale Marine Protected Areas (LSMPAs; areas larger than 30,000 km\textsuperscript{2} \emph{sensu} \cite{desanto_2013}) represent at least 80\% of the protected area in the ocean (Fig. \ref{fig:LSMPAs_map}; \cite{toonen_2013}). However, very little is known about their human dimensions and implications for fisheries \cite{gray_2017}. Furthermore, research on LSMPAs has focused on their potential ecological benefits, but have left aside the economic implications. One issue of particular importance is that of the displacement or redistribution of fishing effort, which may influence the outcomes of a spatial closure and represent large opportunity costs \cite{smith_2003,smith_2010}.

Displacement or elimination of fishing effort within the protected area might come at a cost to the implementing party when fisheries are managed under an effort market (i.e. fishing rights are sold to firms). For example, tuna purse seine fisheries in the Pacific are collectively managed under a Vessel-Day Scheme (VDS) by nine countries commonly referred to as the Parties to the Nauru Agreement (PNA; Figure \ref{fig:PNA_map}). The agreement regulates the access of foreign fishing vessels to PNA waters, and control the allocation and sales of 45,000 ``vessel-days'': the right for one vessel to fish for one day in a country's waters. Holding 80\% of historical skipjack tuna purse seine grounds within their joint Exclusive Economic Zones (EEZ; with surface area of  14.6 million km\textsuperscript{2}), PNA countries have achieved greater bargaining power by working together \cite{havice_2010}; the vessel-day price rose from \$5,000 USD in 2012 to at least \$9,000 USD in 2016, driven by increasing price floors. The revenue from fisheries access fees represents up to 50\% of government revenue for some of the PNA countries.

The Phoenix Islands Protected Area (PIPA) in Kiribati is one of the most notable Large-Scale Marine Protected Areas on earth. Implemented on January 1st of 2015, PIPA closed an area of 397,447 km\textsuperscript{2}, about 11\% of Kiribati's EEZ or 2.7\% of total PNA area, effectively displacing all fishing effort within it \cite{mcdermott_2018,mccauley_2016}. A pressing question is that of the costs associated with the displacement of fishing effort and possible revenue losses from vessel-day sales. While no studies have assessed the implications of PIPA, other PNA members have pledged the implementation of LSMPAs, like Palau's 80\% closure to go into effect in 2020 \cite{cimino_2019}.

We develop a bioeconomic model of a fishery operating under a vessel-day scheme with spatial closures to explore the effects of effort redistribution resulting from MPA implementation. We characterize the likely outcomes with and without vessel-day trading between countries, as well as the effect of different allocation rules. We then empirically evaluate the behavioral responses and spatial redistribution of the industrial tuna purse seine fleet operating within PNA waters after the implementation of the Phoenix Islands Protected Area, and quantify its economic ramifications and implications to Kiribati. We use the same data and insights from our model to hypothesize what might be the impacts of the proposed Palau National Marine Sanctuary. These are two of the largest protected areas on the planet and both are controlled by PNA countries.

\section{Results}\label{results}

\subsection{Fishing effort markets and marine conservation}

We begin by simulating a spatial closure in a setting where vessel-days can not be traded between countries. A spatial closure in Patch 1 always results in a loss or no-change in revenue from vessel-days, even when the stock moves freely between the protected and non-protected portions of the Patch (Fig. \ref{fig:PNA_model}A). The losses are caused by a drop in vessel-day prices for Patch 1, which results from a lower portion of the stock being available for harvest (Fig. \ref{fig:vessel_day_price_no_trading_plot}). Costs increase with reserve size ($R$), and decrease as within-patch movement ($\theta$) increases. Given the non-linear costs ($\beta > 1$), the closure-to-cost ratio can be greater than 1:1 if within-patch movement is low (\emph{i.e.} $\theta < 0.2$). The spatial closure in Patch 1 results in moderate increases in revenue for Patches 2 - 9 (Fig. \ref{fig:profits_PNA_notKIR_no_trading_plot}). When countries can not trade, the costs of conservation are large and incurred by Patch 1, but the benefits are received by the eight other patches. Importantly, the increases in revenue to the other patches do not compensate for the total losses to Patch 1.

The objective of a VDS is to allow for trading between parties and achieve the optimal distribution of effort in relation to fish distribution. In the case of the PNA, for example, spatial distribution and longitudinal shifts of tuna -particularly skipjack- have been linked to ENSO events \cite{lehodey_1997}. Vessels redistribute across PNA countries as they track fish, and tradable permits reduce costs to the parties by allowing them to sell surplus vessel-days to be used in the EEZ of another country \cite{aqorau_2018}. We explore the effect of a spatial closure in Patch 1, but now allow for vessel-days to be traded between Patches. Patch-level demand for vessel days is governed by the fishable stock size (\emph{i.e.} biomass not within the reserve). Trading occurs until vessel-day prices are the same across patches, given a total allowable effort (\emph{i.e.} 45,000 vessel-days) is reached. Here, costs from a closure in Patch 1 are significantly reduced by trading (Fig. \ref{fig:PNA_model}B-C). In this case, the surplus of vessel-days in Patch 1 are sold to Patches 2 - 9. Vessel-day prices are constant across patches and lower than when there is no reserve, but still higher than when no trading is allowed (Fig. \ref{fig:vessel_day_price_with_trading_plot}). By optimally trading vessel-days, the cost to the implementing country can be reduced as much as 97\%.

Our previous analysis predicts a loss in total revenue from effort displacement. Annual vessel-days are often allocated based on a combination of historical within-patch effort and biomass. In the PNA, for example, sixty percent of the allocation is calculated based on EEZ effort over the last seven years and 40\% is calculated based on the 10-year average of each country’s share of estimated skipjack and yellowfin biomass within its EEZ.\footnote{This is explained in more detail in Article 12.5 of the 2012 Amendment to the Palau Agreement and in \cite{Hagrannsoknir2014}.} Trading vessel-days to other countries would imply that historical within-patch effort declines through time. The allocated days to a patch with a full spatial closure would eventually be reduced to just the 40\% based on biomass.

We explore the effect that allocation rules have through time when trading is allowed by projecting the fishery 50 years into the future. We used a range of allocation rules that weight effort ($\alpha$) and biomass ($1 - \alpha$) differently to update annual allocations based on a seven-year running mean of these measures. Costs are greater when allocation is based only on patch-level historical effort (\emph{i.e.} $\alpha = 1$; Fig. \ref{fig:allocation_cost_plot}). However, costs can be significantly reduced when trading is allowed if allocation is based on historical biomass instead. When there is no trading, patch-level effort remains unchanged, and since any growth in stock is equally distributed across patches, allocation rules do not affect costs.

\subsection{Phoenix Islands Protected Area}

We use identification of fishing activity via Automatic Identification Systems (AIS) to track 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018 \cite{mccauley_2016,kroodsma_2018}. We track the spatial redistribution of the fleet and explore behavioral changes following the 2015 implementation of PIPA in Kiribati. We continuously observe 92 vessels for the 2012 - 2018 period. Of these, 64 vessels fished within PIPA at least once prior to its implementation, and we refer to them as the ``displaced vessels''. The remaining 28 vessels never fished in PIPA waters, and we refer to them as ``non-displaced vessels''. The group with the remaining 221 vessels contains vessels that were not continuously observed before and after the implementation of PIPA, and we refer to these as ``other vessels''. Our analyses incorporate the monthly NINO4 anomaly index (Fig. \ref{fig:nino_plot}) to control for the association that fish and vessels might have to environmental variation \cite{lehodey_1997,kroodsma_2018,aqorau_2018}.

A market for fishing effort and allocation rules can reduce the costs of displacing fishing effort, but a spatial closure is still likely to cause behavioral changes as vessels redistribute. We find that between 2015 and 2016, displaced vessels spent 3,916 and 2,249 less vessel-days in Kiribati and PNA waters, respectively (Figs. \ref{fig:all_PS_VDS_year}). Over the same period, non-displaced and other vessels spent 1,278 less days in Kiribati, but spent 9,853 more days in PNA waters overall. These result in a net loss of 5,195 vessel-days for Kiribati, and a net gain of 7,600 vessel-days at the PNA-level (\emph{i.e.} the other 8 countries). At a vessel-day price of \$9,000 USD per day, these would represent \$46.7 and \$68 million USD, respectively.

Annual vessel-days in Kiribati continued to decrease to just 7,479 in 2018, mainly due to displaced vessels allocating less time to Kiribati (Fig. \ref{fig:all_PS_VDS_year}B, Fig. \ref{fig:hist_kir_fishing}). Current effort in Kiribati waters is well below their 11,000 vessel-day allocation \cite{yeeting2018stabilising}. However, we observe a relatively constant 45,000 vessel-days at the PNA-level, suggesting that Kiribati is, in fact, able to trade vessel-days and reduce the costs of displacement. The spatial redistribution patterns of displaced vessels relative to non-displaced vessels are shown in Figure \ref{fig:fishing_raster_diff}.

We complement our analysis of change in observed vessel-days by looking at country-level data reported by the Pacific Islands Forum Fisheries Agency (FFA). We find that Kiribati's revenue went from \$148.8 million USD in 2015 to \$118.3 million USD in 2016, representing a decrease of \$30.5 million USD. Total PNA revenues showed a net increase of \$28 million USD (Fig. \ref{fig:total_PNA_revenues}). The largest decrease was observed for Kiribati, while the largest increase was observed for Papua New Guinea.

Total fishing effort in the PNA increased between 2015 and 2016, but fishing area decreased due to PIPA implementation (Fig. \ref{fig:all_PS_VDS_year}). We inspect the crowding effects that may arise by applying more (or the same amount of) fishing effort over less fishing area with two indices of spatial overlap between displaced and non-displaced vessels: 1) the number of cells that had fishing activity from both groups for each month and 2) the correlation of presence/absence of fishing activity between both groups over one month. The two fleets significantly interact more with each other after the implementation of PIPA (Table \ref{tab:sp_corr}, Fig. \ref{fig:sp_corr}). The number of cells with presence from both fleets and spatial correlation increase by a factor of four and three, respectively. Crowding might be driven by environmental variation concentrating fish at specific locations, but the NINO4 anomaly alone explained just 3\% and 7\% of the variation in our crowding measures, compared to the 70\% variation explained when accounting for PIPA implementaiton. We observe similar patterns when repeating the crowding exercise for Kiribati's EEZ only (Table \ref{tab:KIR_sp_corr}, Fig \ref{fig:KIR_sp_corr}).

Effort displacement and crowding are likely to induce behavioral changes in the fishing fleet. We calculate nine vessel-level measures that could capture responses to spatial closures: daily fishing hours, daily non-fishing at-sea hours, the proportion of fishing to non-fishing at-sea hours, daily distance traveled, daily mean distance from shore of fishing events (km), daily mean distance from port of fishing events (km), as well as monthly hours spent in PNA waters, Kiribati waters, and the High Seas (Fig. \ref{fig:all_panels}). We find no evidence of displaced vessels fishing more after PIPA implementation, and in fact observe a 27.5\% decrease ($p < 0.01$; Table \ref{tab:main_DID}) relative to non-displaced vessels. Likewise, we observe a 3.4\% decrease of fishing hours relative to total at-sea hours ($p < 0.01$; Table \ref{tab:main_DID}). Treated vessels traveled 23.3\% less distance, and fishing events occurred 32.9\% and 16.9\% closer to shore and to port, respectively. These changes in distance from shore and port are likely caused by redistribution, as we observe that displaced vessels fished 79.67\% and 49.6\% less in Kiribati and PNA waters respectively, compared to the trend observed for non-displaced vessels ($p < 0.01$). We do not observe a statistically significant increase in non-fishing at-sea hours (\emph{i.e.} a proxy for search time) and fishing hours on the High Seas. We repeat this analysis for sample restrictions where we exclude all Chinese vessels (Table \ref{tab:DID_without_CHN}), all PNA-owned vessels (Table \ref{tab:DID_without_PNA}) and all Taiwanese and USA vessels (Table \ref{tab:DID_without_USA_TWN}) and find qualitatively the same responses.

\subsection{Palau National Marine Sanctuary}

On October 28, 2015, the President of Palau signed into law the Palau National Marine Sanctuary (PNMS) Act. Starting in December 2020, this Act will close 500,000 km\textsuperscript{2} (about 80 percent of Palau’s EEZ) to commercial fishing activities, creating the 14th largest protected area in the world, effectively displacing all effort within its waters (Fig. \ref{fig:plw_2018} shows longline and purse seine fishing effort within PNMS for 2018). As part of the Nauru Agreement, Palau receives 700 purse seine vessel-days each year. They also  use a vessel-day scheme to manage the longline fishery  (10,500 vessel-days per year). Unlike the purse seine vessel-days, longline vessel-days are not tradable across countries and are worth much less, at about \$200 USD each. Our previous analyses suggest that Palau should be able to lessen the costs of the closure by trading purse seine vessel-days (\$5.65 - \$8.75 million USD) with other PNA members, but might lose more than 50\% of the revenues from the longline vessel-days (\$2.1 million USD).

Table \ref{tab:revenue_loss} presents estimates of the potential revenue losses following full enactment of the PNMS under four different scenarios. In Scenario 1, Palau retains its current allotment of purse seine vessel-days and is able to sell them for a similar price to what it is currently selling them to the United States under the South Pacific Tuna Treaty (also known as the Multilateral Treaty on Fisheries; \$12,500/day). In Scenario 2, Palau is able to keep its current allotment of purse seine VDS to transfer to other PNA countries at the current benchmark price (\$8,000/day). Scenario 2 is likely if Palau retains its current allocation, but the US no longer purchases days. It should be noted that if allocation continues to be calculated based on effort and biomass, and if Palau continues to be allocated vessel days, its allocation will decrease as purse seine effort in its EEZ eventually reaches zero. In Scenarios 3 and 4, Palau loses all of its purse seine VDS, at \$8,000/day and \$12,500/day, respectively. In all scenarios, all longline vessel days and export tax revenues are lost, since longline vessel days are currently not tradable and Palau is currently planning on banning the export of fish. The longline vessel day loss is calculated using an average value of \$200 for 10,500 days. The export tax loss is calculated given the average tax revenue from 2012-2014 (\$482,236  from \cite{Gillett2016}).

% I think we need to modify the last  part of this. Based on our analyses, we can say that they would lose a large portion of the longline days, but not all. I can calculate the longline effort within the PNMS and we can describe a ``worst-case scenario''. Turns out that the 80% closure has only 55.75% of all fishing effort. Also, my understanding from the workshop is that the export ban is already in place EXCEPT for tuna purse seines. They also pay a 0.35 USD / Kg export tax for fish LANDED in Palau. The new proposal is to LIFT the export ban and push for the development of a domestic pelagic fishery (artisanal mFADs and Pole and Line).

\input{revenue_loss}

\section{Discussion}

Our findings provide insights into the effect that LSMPAs can have on the redistribution of fishing effort and changes in behavior. Our analyses predict losses in revenue to countries that implement a spatial closure. However, these losses can be significantly reduced by market instruments and biomass-based allocation rules. Using vessel tracking data, we see that vessels that fished inside PIPA before the implementation redistribute to other areas within Kiribati, but also to other PNA countries and the high seas. Surprisingly, there is no drop off in fishing effort in Kiribati in 2015 but a noticeable drop from 2016 onward. Redistribution leads to a crowding effect, and the spatial closure has little effect on the \emph{total} fishing effort exerted by purse seiners, but vessel-level responses were heterogeneous. Our analysis suggests that losses would be expected for Palau's Marine Sanctuary, but that these can be minimized if effort is traded freely and allocations are based on biomass, not effort. Here, we discuss the implications of our findings and possible shortcomings in our analysis.

>>>>

\title{Fishery Markets and Large-Scale Marine Conservation}

@@

\author{Juan Carlos Villase\~{n}or-Derbez,$^{1\ast}$ John Lynham,$^{2}$ Christopher Costello$^{1}$\\
\\
\normalsize{$^{1}$Bren School of Environmental Science \& Management,}\\
\normalsize{University of California at Santa Barbara, Santa Barbara, CA}\\
\normalsize{$^{2}$Department of Economics, University of Hawaii at Manoa, Honolulu, HI}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail: juancarlos@ucsb.edu.}
}

@@

% Place your abstract within the special {sciabstract} environment.
\begin{sciabstract}
Will rights-based approaches to managing natural resources facilitate, or impede large-scale conservation? We study this question in the context of large-scale marine protected areas, and ask whether a country that commits to large-scale conservation is destined to suffer commensurate losses in fishery income.  Indeed, this is a question maritime nations all over the world face as they consider whether, and how, to comply with international commitments to expand Marine Protected Area (MPA) coverage. We use a combination of spatial bioeconomic modeling, which accounts for fisheries markets such as the vessel-day trading scheme in the Pacific Ocean, and spatially-explicit vessel-level empirical data on fishing behavior before and after MPA implementation, to assess how market design incentivizes, or penalizes conservation. The answer hinges critically on two design features of the rights-based fishery market. First, only if trading of harvest or vessel-days is allowed across countries can a conservation-minded country capture the economic spillover benefits of their conservation actions.  Instead, if trading is forbidden, countries are likely to suffer losses from conservation. Second, when rights are allocated between countries, the allocation must depend only on biological features, and not on economic factors such as a country's past effort or catch. Otherwise, the allocation process acts as a punishment to the conserving country. Overall, these results show that seemingly benign design features of a market can have indelible impacts on a country's willingness to engage in large-scale conservation; these results are confirmed with an empirical case study from the Phoenix Islands Protected Area, one of the world's largest marine reserves. With the right design features, large-scale protected areas such as the Phoenix Islands can be achieved at close to zero cost to the implementing country.
\end{sciabstract}

@@

%%%% Other titles %%%%%%%%%%%%
% Fishery markets enable marine conservation
% Designing environmental markets for marine conservation
% Large-scale marine conservation is enabled by market instruments
% Rights-based management to promote marine conservation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Recognizing a need to secure biodiversity, blue carbon, rare habitats, and other ecosystem services, various international bodies have committed to dramatically expand marine protection around the world \cite{dinerstein_2019}. Currently only about 3\% of the world's ocean is strongly protected \cite{sala_2018}, and these commitments call for 10\%, 30\%, or even up to 50\% of the ocean to be off-limits to fishing. Fundamentally, to achieve these goals, huge swaths of sovereign nations' waters will have to be closed to fishing. What incentives might motivate a country to engage in such large-scale marine protection? Would any country rationally close 20\%, 50\%, or even 100\% of its waters to fishing if this meant losing all fishing revenue from within its borders? We show that a country's incentives for large-scale conservation hinge critically on an unsuspected culprit: the design of fishing rights.

We are motivated by a real-world institution called the Parties to the Nauru Agreement (PNA). Like an OPEC for tuna, the PNA is a nine-country coalition of Pacific Island states that collectively manages tuna fishing in its members' waters\cite{havice_2013,aqorau_2018}. These waters collectively account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental United States), and over 60\% of skipjack tuna catches in the Western Central Pacific \cite{havice_2013}.

In addition to highly productive tuna fisheries, the PNA waters provide a wealth of other ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. The PNA member nations derive enormous benefits from leasing tuna fishing rights to foreign fleets, in some cases exceeding half of a country's GDP. A natural concern arises: even if the closure creates net benefits to the world, and perhaps even to tuna fisheries Pacific-wide, the economic loss to the conserving country could still be extremely large, and may prevent it from engaging in any meaningful conservation. Intuitively, then, it seems that a country's willingness to engage in large-scale conservation will depend crucially on whether it can successfully capture a sufficiently high fraction of the benefits that are derived from that conservation. We show how a rights-based fisheries market can be designed (if it is new) or modified (if it already exists) to ensure that the conserving country is rewarded, not punished, for large-scale conservation.

To mirror the strategies and spatial connections among the nine PNA countries and the high seas, we develop a 10-patch spatial bioeconomic model. The fishery in nine of the patches operates under a market-based system in which ``vessel days'' are capped, closely tracked, and may be traded across countries.  This vessel-day scheme (VDS) is actually used by the PNA; each country is given an allocation of vessel-days that can then be sold to fishing firms. A vessel-day owned by country $i$ entitles $i$ to lease that day to a fishing vessel, entitling it to fish in $i$'s EEZ for a 24 hour period. In our model, Patches 1 - 9 are countries that operate under a VDS (in which fishing days are limited) and Patch 10 represents the high seas, where fishing days are governed by economic conditions, not by the VDS.

While modeling the VDS market may be interesting in its own right, our primary purpose is to examine how the VDS affects a country's incentives for large-scale conservation. To examine this, suppose that Country 1 considers closing a portion $R$ (between 0 and 1) of its waters as a marine reserve, in which no fishing is allowed. When Country 1 creates a reserve, several consequences may arise.  Perhaps most obviously, the area available to fish in Country 1 is now limited, which could reduce the amount of money fishing firms are willing to pay for a day of fishing in $i$'s waters.  Second, to the extent that catch is reduced in Country 1, fish may become more abundant elsewhere; this could imply an increase in the lease price to fish in the other eight PNA countries' waters because the VDS price reflects the profit from fishing in a country, which depends crucially on the abundance of fish. For the same reason, the increase in abundance could attract new entrants on the high seas, who could quickly dissipate the benefits there. The purpose of our model is to examine these consequences in detail, with a particular focus on how design features of the market affect Country 1's incentives to create the marine reserve.

Not all market-based approaches to environmental management are created equal. A pervasive finding across a range of natural resources is that features of market design can have enormous implications for the market's functioning. In the context of fishery markets, two market design features stand out as important for a variety of outcomes, and we will show that they are pivotal in determining the incentives for large-scale marine conservation. The first design feature has to do with the trading of rights, and the second design feature concerns the allocation of rights. We examine the effects of large-scale conservation in Country 1 across a range of scenarios for trading and allocation of vessel day fishing rights across our nine model countries. In all cases we solve for the equilibrium vessel-day price, fishing effort redistribution, and fish stock that would be expected to occur in the market. We quantify the change in revenue to each PNA country from large-scale conservation in Country 1, by comparing revenues from each scenario to a benchmark scenario without any conservation action. We simulate these outcomes across a range of reserve sizes and assumptions about within-patch stock movement.

Why would trading and allocation rules affect the incentives for conservation by Country 1?  Invoking \emph{reductio ad absurdum}, consider the incentives for Country 1 to close 100\% of its waters. Such a closure would surely benefit the other countries through the increased abundance of fish.  If Country 1 could sell its fishing rights, the other countries would likely buy them, and this could offset the foregone costs of fishing its own waters. But if Country 1 was not allowed to sell its rights, then it would lose all of its fishing revenue.  This provides the simple intuition for why allowing trading across countries is so critical to conservation incentives. The rules for how fishing rights are allocated across the nine countries are equally important. Suppose rights are allocated each year based on the previous year's fishing - the more a country fishes, the more it gets allocated the next year. This would clearly disadvantage a conservation-minded country and, in fact, might reward undesirable behavior.

In our bioeconomic model, we examine the implications of conservation under alternative market designs: with, and without trading, and under alternative allocation rules. To test how the PNA allocates rights, we vary the degree to which periodic allocation is based on historical within-patch effort versus the actual abundance of fish biomass in a patch. Finally, we use vessel-tracking data and official reports of government revenue to corroborate our results by looking at the Phoenix Islands Protected Area, implemented in 2015 by Kiribati within the PNA.

We begin by simulating a fishery where trading of fishing days across countries is forbidden. In that setting, we find that a spatial closure in Patch 1 will always result in a loss in revenues to Patch 1 (Fig. \ref{fig:PNA_model}A). Closing a greater proportion of the patch results in greater costs, but these decrease if within-patch stock movement ($\theta$) increases, allowing vessels to harvest the stock within the remaining open area. The closure-to-cost ratio for any reserve size is greater than 1:1 when stock movement is low (\emph{i.e.} $\theta < 0.2$), implying that a 30\% closure would result in at least a 30\% loss in revenues. Even for a highly mobile stock, a reserve reduces the amount of biomass that is available for harvest in Patch 1, which reduces a vessel's willingness to pay to fish in such waters (Fig. \ref{fig:vessel_day_price_no_trading_plot}). When countries cannot trade, the costs of conservation are large and incurred by Patch 1, but the benefits are received by the eight other patches (Figure S2 shows that the increase in revenue ranges from 0\% to 6\%).

How would a different market design, where trading between countries is allowed, change these results? To answer this we simulate the exact same fishery, but now allow for vessel-days to be traded across patches. A closure in Patch 1 lowers the value of vessel-days in that patch. Vessel-days in patches 2 - 9 are now worth more causing vessel-days from Patch 1 to be traded to Patches 2 - 9. Trading continues until vessel-day prices are the same across all patches (Fig. \ref{fig:vessel_day_price_with_trading_plot}). Under this market design, we find that revenue losses are less than 1\%, compared to the benchmark scenario with no reserve (Fig. \ref{fig:PNA_model}). Overall, we find that between 88\% and 99\% of the costs of conservation can be avoided if markets are designed to enable trading (Fig. \ref{fig:PNA_model}C).

To analyze the consequences of different allocation rules on conservation incentives, we simulate an annual allocation that is a weighted function of historical fishing in a country and historical fish biomass in a country. If allocation rules are based only on the historical effort in a country, then trading vessel-days to other countries would imply a future reduction in allocation. We simulate the fishery 50 years into the future and annually re-allocate vessel-days based on a 7-year running mean of patch-level effort and biomass. We use a range of allocation rules to weight effort ($\alpha$) and biomass ($1 - \alpha$) and compare the revenues to a fishery operating for 50 years without any closures. We find that when allocation is based on historical effort only (\emph{i.e.} $\alpha = 1$), implementing a reserve results in losses of 20\% to 93\%, depending on the size of reserve and stock movement (Fig. \ref{fig:allocation_cost_plot}). However, these costs are reduced as more weight is put on historical biomass. A biomass-only allocation rule (\emph{i.e.} $\alpha = 0$) results in costs as low as 0.1\% to 0.7\%, essentially eliminating the costs of conservation for all reserve sizes and possible movement. This result implies that if allocation is based purely on the biological features of a stock (\emph{i.e.} the biomass within a country's waters), and not on fishing effort, the incentives for conservation are much greater.

We have shown that design features of a rights-based fisheries management system have enormous implications for the incentives a country faces to engage in large-scale conservation. Allowing trading of rights across countries, and re-allocating annual rights based on biological variables, not fishing pressure, can essentially eliminate the cost of conservation, even for very large reserves. We now move to an empirical setting where we combine vessel-tracking data and the implementation of one of the world's largest marine reserves to identify the effort displacement and likely costs of conservation.

The Pacific island nation of Kiribati implemented the Phoenix Islands Protected Area in January 2015 (Fig. \ref{fig:PNA_map}). With an extension of 397,447 Km\textsuperscript{2}, this 11.5\% closure effectively displaced all fishing effort within its boundaries \cite{mccauley_2016,mcdermott_2018}. We use identification of fishing activity via Automatic Identification Systems (AIS) provided by Global Fishing Watch to track 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018 \cite{kroodsma_2018}. We continuously observe 92 vessels for the 2012 - 2018 period. Of these, 64 vessels fished within PIPA at least once prior to its implementation, and we refer to them as the ``displaced vessels''. The remaining 28 vessels never fished in PIPA waters, and we refer to them as ``non-displaced vessels''. The group with the remaining 221 vessels contains vessels that were not continuously observed before and after the implementation of PIPA, and we refer to these as ``other vessels''.

We use the tracking data to calculate the time that vessels spent within Kiribati and PNA waters (Figs. \ref{fig:all_PS_VDS_year}A-B). Our observations match the 45,000 vessel-days allowed at the PNA level, and roughly match the 11,000 vessel-days historically allocated to Kiribati \cite{yeeting2018stabilising}. We find that displaced vessels fished 922 less days (a 10\% reduction) in Kiribati between 2014 and 2015, immediately after PIPA was implemented. For the same period, these vessels spent just 48 days less in PNA waters, suggesting that trading might have facilitated relocation of effort within PNA waters. On the other hand, non-displaced and other vessels spent 1,621 and 2,034 additional days in Kiribati in 2015, relative to 2014. The same pattern is observed at the PNA level, with vessels spending an additional 3,789 and 3,691 vessel-days, respectively. These changes result in a net decrease of fishing effort within Kiribati, from 12,282 in 2014 to just 7,542 in 2018, with displaced vessels driving the decrease (Figs. \ref{fig:all_PS_VDS_year}A). The spatial redistribution patterns of displaced vessels relative to non-displaced vessels suggest that some relocated to other waters in Kiribati (\emph{i.e.} the Gilbert islands and Line islands), but also the Marshall Islands and the high seas (Fig \ref{fig:fishing_raster_diff}).

Our analysis suggests that the implementation of PIPA resulted in a decrease in fishing effort within Kiribati's waters. But does this result in losses to Kiribati? We do not explicitly observe trading between countries, but combining different sources of information we are able to paint a comprehensive picture. We use country-level fisheries revenue data reported by the Pacific Islands Forum Fisheries Agency (FFA) for the 2008 - 2016 period to complement our analysis (Fig. \ref{fig:revenues}A). Kiribati's reported revenue increased from \$127.3 million USD in 2014 to \$148.8 million USD in 2015, before decreasing to \$118.3 million USD in 2016. This increase and subsequent decrease in revenues matches the AIS patterns we observe for Kiribati in 2014 - 2016 (Fig. \ref{fig:all_PS_VDS_year}A), and observed AIS effort is a reliable predictor of government revenue across all PNA countries from 2008 to 2016 (Fig. \ref{fig:revenue_FFA_effort_GFW}). At the PNA level, total revenues showed a net increase of \$64.7 and \$28 million USD for 2015 and 2016 respectively (Fig. \ref{fig:total_PNA_revenues}).

\textbf{We also inspect the crowding effects that may arise by applying more (or the same) fishing effort over less fishing area by inspecting the spatial overlap between displaced and non-displaced vessels over a gridded surface(Fig \ref{fig:fishing_raster_diff}). These groups interacted more with each other after the implementation of PIPA (Table \ref{tab:sp_corr}, Fig. \ref{fig:sp_corr}). The number of cells with presence from both fleets and the spatial correlation between groups increases by a factor of four and three, respectively. Environmental variation may certainly drive this, but the NINO4 anomaly \ref{fig:nino_plot} alone explained just 3\% and 7\% of the variation in our crowding measures, compared to the 70\% variation explained when accounting for PIPA implementation. We observe similar patterns when replicating the crowding exercise for Kiribati's EEZ only (Table \ref{tab:KIR_sp_corr}, Fig \ref{fig:KIR_sp_corr}).}

\textbf{We use multiple linear regression to measure the change in nine vessel-level measures for displaced and non-displaced vessels before and after the implementation of PIPA (Fig. \ref{fig:all_panels}; Table \ref{tab:main_DID})). We find no evidence of displaced vessels fishing more after the implementation, and in fact observe a 27.5\% decrease relative to non-displaced vessels ($p < 0.01$). Likewise, we observe a 3.4\% decrease of fishing hours relative to total at-sea hours ($p < 0.01$). Displaced vessels traveled 23.3\% less distance, and fishing events occurred 32.9\% and 16.9\% closer to shore and to port, respectively. These changes in distance from shore and port are likely explained by redistribution, as we observe that displaced vessels fished less in Kiribati and PNA waters, compared to the trend observed for non-displaced vessels ($p < 0.01$). We do not observe changes in non-fishing at-sea hours (\emph{i.e.} a proxy for search time) and fishing hours on the High Seas. We repeat this analysis for sample restrictions where we exclude all Chinese vessels (Table \ref{tab:DID_without_CHN}), all PNA-owned vessels (Table \ref{tab:DID_without_PNA}) and all Taiwanese and USA vessels (Table \ref{tab:DID_without_USA_TWN}) and find qualitatively the same responses.}

A new large-scale marine protected area is being proposed in the PNA. On October 28, 2015, the President of Palau signed into law the Palau National Marine Sanctuary (PNMS) Act. Starting in December 2020, this Act will close 500,000 km\textsuperscript{2} (about 80\% of Palau's EEZ) to commercial fishing activities \cite{cimino_2019}, creating the 14\textsuperscript{th} largest protected area in the world (Fig. \ref{fig:PNA_map}). As a PNA member, Palau receives 700 purse seine vessel-days each year. Their longline fishery is also managed under a vessel-day scheme, with 10,500 non-tradable vessel-days per year, which currently sell at about \$200 USD each (Fig. \ref{fig:PLW_GFW_ts}). What does an 80\% closure represent for Palau's revenue from access fees? AIS data (2012 - 2019) shows that PNMS would displace $65.6 \pm 0.08$\% and $82.2 \pm 0.08$\% of longline and purse seine activity (Figs. \ref{fig:plw_2018} and \ref{fig:plw_ts_plot}). Palau may be able to reduce the cost of lost license revenue if purse seine vessel-days are traded with other PNA members. However, the current reallocation rule for purse seine vessel-days is based on historical effort and biomass, which would lead to a decreased allocation through time. Palau, howver, might continue to sell purse seine vessel-days to the United States under the South Pacific Tuna Treaty (also known as the Multilateral Treaty on Fisheries; \$12,500/day). Longline vessel days are currently not tradable, meaning that the price for longline vessel-days might decrease, leading to losses in profits.
